---
title: "Movie ratings and box office performance"
author: "IM Lahart"
date: "2024-03-13"
output: html_document
---

### Page set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Install renv for reproducibility, include=FALSE}
#Install renv package if needed
if(!require(renv)){install.packages('renv')}
```

```{r initializing renv, include=FALSE}
renv::init()
```


### Install and load packages

```{r Install and load packages, include=FALSE}
#Install tidyverse package if needed
if(!require(tidyverse)){install.packages('tidyverse')}
#Load tidyverse package
library(tidyverse)

#Install tables package if needed
if(!require(tables)){install.packages('tables')}
#Load tables package
library(tables)

#Install flextable package if needed
if(!require(flextable)){install.packages('flextable')}
#Load flextable package (as_flextable)
library(flextable)

#Install arsenal package if needed
if(!require(arsenal)){install.packages('arsenal')}
#Load arsenal package (for tableby)
library(arsenal)

#Install modelsummary package if needed
if(!require(modelsummary)){install.packages('modelsummary')}
#Load modelsummary package (for datasummary_crosstab)
library(modelsummary)

#Install RColorBrewer package if needed
if(!require(RColorBrewer)){install.packages('RColorBrewer')}
#Load RColorBrewer package
library(RColorBrewer)

renv::snapshot()
```

### Read file in

```{r Read in csv file, include=FALSE}
Movie_db <- read_csv("Movie_database.csv", skip_empty_rows = TRUE)
```

### Processing of data
```{r Processing variables, include=FALSE}
Movie_db$Genre <- as.factor(Movie_db$Genre)
Movie_db$Best_picture <- as.factor(Movie_db$Best_picture)
Movie_db$Genre <- fct_collapse(Movie_db$Genre, Crime_mystery = c("Crime", "Mystery"))
# head(Movie_db) # If you would like to check your dataset, just delete the hashtag before 'head' at the beginning of this line.
```


```{r Create dataframe for tables, include=FALSE}
Movies_genre <- Movie_db %>% dplyr::select(c(Genre, IMDB_Rating, MetaCritic, Budget_m, Runtime, Worldwide_earning_m, Best_picture))
```

### Summary table using tably and summary

```{r Table, results="asis"}
mylabels <- list(Best_picture = "Best picture", MetaCritic = "Meta critic score", IMDB_rating = "IMDB rating", Runtime = "Run time (mins)", Budget_m ="Budget, million $", Worldwide_earning_m = "Global earnings, million $")
tab1 <- tableby(Genre ~ Best_picture + IMDB_Rating + MetaCritic + Runtime + Budget_m + Worldwide_earning_m, data=Movie_db, test = FALSE)
summary(tab1, labelTranslations = mylabels, digits = 1)
```

### Table 1. Movie viewer and critic ratings, runtime, budget, and earnings per genre

```{r Genre Table}
set_flextable_defaults(background.color = "white")

tab <- tabular(
    (Genre + 1) ~ (n = 1) + Format(digits = 1) *
      (IMDB_Rating + MetaCritic + Runtime + Budget_m + Worldwide_earning_m) * (mean + sd),
    data = Movies_genre)
  
Genre_tab <- as_flextable(tab) %>%
            autofit()

Genre_tab <- theme_vanilla(Genre_tab)
Genre_tab <- add_footer_lines(Genre_tab, "Note: IMDB rating 0-10; MetaCritic score 0-100%; Budget and Worldwide earnings in millions of dollars")
Genre_tab <- color(Genre_tab, part = "footer", color = "#666666")
Genre_tab <- set_caption(Genre_tab, caption = "Table 1. Movie viewer and critic ratings, runtime, budget, and earnings per genre")
Genre_tab
```


```{r Table 1 dimensions}
dims <- flextable_dim(Genre_tab)
dims
```

```{r Table 1 save, fig.height=8.1, fig.width=4.1}
save_as_image(Genre_tab, path = "Movie data by genre.png")
```


### Table 2. Movie viewer and critic ratings, runtime, budget, and earnings by best picture nomination

```{r Best Picture Table}
set_flextable_defaults(background.color = "white") 

tab_bp <- tabular(
    (Best_picture + 1) ~ (n = 1) + Format(digits = 1) *
      (IMDB_Rating + MetaCritic + Runtime + Budget_m + Worldwide_earning_m) * (mean + sd),
    data = Movies_genre)
  
BP_tab <- as_flextable(tab_bp) %>%
            autofit()

BP_tab <- theme_vanilla(BP_tab)
BP_tab <- add_footer_lines(BP_tab, "Note: IMDB rating 0-10; MetaCritic score 0-100%; Budget and Worldwide earnings in millions of dollars")
BP_tab <- color(BP_tab, part = "footer", color = "#666666")
BP_tab <- set_caption(BP_tab, caption = "Table 2. Movie viewer and critic ratings, runtime, budget, and earnings by best picture nomination")
BP_tab
```

```{r Table 2 dimensions}
dims <- flextable_dim(BP_tab)
dims
```

```{r Table 2 save, fig.height=7.9, fig.width=2.1}
save_as_image(BP_tab, path = "Movie data by best picture.png")
```


### Table 3. Best picture nominations by genre

```{r Table of best picture nominations by genre}
options("modelsummary_factory_default" = "flextable")
Genre_ctabs <- datasummary_crosstab(Best_picture ~ Genre, data = Movie_db)

Genre_ctabs <- theme_vanilla(Genre_ctabs)
Genre_ctabs <- align(Genre_ctabs, align = "center", part = "all")
Genre_ctabs <- autofit(Genre_ctabs)
Genre_ctabs <- add_footer_lines(Genre_ctabs, "Note: Yes = movies that were nominated for best picture at the academy awards; No = movies that were not nominated.")
Genre_ctabs <- color(Genre_ctabs, part = "footer", color = "#666666")
Genre_ctabs <- set_caption(Genre_ctabs, caption = "Table 3. Best picture nominations by genre")
Genre_ctabs
```

```{r Table 3 dimensions}
dims <- flextable_dim(Genre_ctabs)
dims
```

```{r Table 3 save, fig.height=7.5, fig.width=1.75}
save_as_image(Genre_ctabs, path = "Best picture nominations by movie genre.png")
```


### Figure 1. The relationship between critics opinions and how much a movie earns

```{r Genre plot}
Plot_1 <- ggplot(Movie_db, aes(x=MetaCritic, y=Worldwide_earning_m, color = Genre)) + 
          geom_point() +
          theme_classic()

Genre_plot <- Plot_1 + ggtitle("The relationship between critics opinions and how much a movie earns") +
                      xlab("Meta Critic Score") + 
                      ylab("Global Earnings in millions ($)") + 
                      scale_color_brewer(palette = "Set1")

Genre_plot
```

```{r Genre plot save}
ggsave("Critic ratings and movie earnings.png", plot = Genre_plot, width = 9.0, height = 6.0, dpi = 300)
```


### Figure 2. The relationship between movie run time and movie earnings

```{r Run time plot}
Plot_2 <- ggplot(Movie_db, aes(x=Runtime, y=Worldwide_earning_m, color = Genre)) + 
          geom_point() +
          theme_classic()

Runtime_plot <- Plot_2 + ggtitle("The relationship between movie run time and movie earnings") +
                          xlab("Run time (mins)") + ylab("Global Earnings in millions ($)") + 
                          scale_color_brewer(palette = "Set1")

Runtime_plot
```

```{r Runtime plot save}
ggsave("Runtime and movie earnings.png", plot = Runtime_plot, width = 9.0, height = 6.0, dpi = 300)
```

### Figure 3. Best picture nominations by genre

```{r Barplot Best picture by genre}
p1 <- ggplot(data = Movie_db, aes(x = Genre, fill = Best_picture)) +
  geom_bar(position = "dodge") +
  scale_y_continuous(limits = c(0, 30)) +
  labs(subtitle = "Best picture nominations by genre") + 
  xlab("Genre") + 
  ylab("N Best picture nominations") +
  theme_classic()
p1
```

```{r Barplot Best picture by genre save}
ggsave("Barplot Best picture by genre.png", plot = p1, width = 7.0, height = 4.0, dpi = 300)
```


### Figure 4. Movie earnings by genre

```{r Boxplot movie earnings by genre}
p <- ggplot(Movie_db, aes(x=Genre, y=Worldwide_earning_m, color = Genre)) + 
  geom_boxplot()

Box_earnings <- p + theme_classic() + 
            xlab("Movie genre") +
            ylab("Worldwide earnings in millions ($)") +
            stat_summary(fun.y=mean, geom="point", shape=23, size=4, color = "red") + 
            scale_fill_brewer(palette="Set2")

Box_earnings
```

```{r Boxplot movie earnings by genre save}
ggsave("Boxplot movie earnings by genre.png", plot = Box_earnings, width = 9.0, height = 6.0, dpi = 300)
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
